---
output:
  md_document:
    variant: markdown_github
html_preview: false
---


```{r echo=FALSE, results="hide", message=FALSE}
library("ggtree") 
library("txtplot")
library("scholar")
library("dlstats")
library("rvcheck") 
get_devel_version <- rvcheck:::get_devel_version
```


# DOSE: Disease Ontology Semantic and Enrichment analysis

[![platform](http://www.bioconductor.org/shields/availability/devel/DOSE.svg)](https://www.bioconductor.org/packages/devel/bioc/html/DOSE.html#archives)
[![Build Status](http://www.bioconductor.org/shields/build/devel/bioc/DOSE.svg)](https://bioconductor.org/checkResults/devel/bioc-LATEST/DOSE/)
[![Linux/Mac Travis Build Status](https://img.shields.io/travis/GuangchuangYu/DOSE/master.svg?label=Mac%20OSX%20%26%20Linux)](https://travis-ci.org/GuangchuangYu/DOSE)
[![AppVeyor Build Status](https://img.shields.io/appveyor/ci/Guangchuangyu/DOSE/master.svg?label=Windows)](https://ci.appveyor.com/project/GuangchuangYu/DOSE)
[![install with bioconda](https://img.shields.io/badge/install%20with-bioconda-green.svg?style=flat)](http://bioconda.github.io/recipes/bioconductor-dose/README.html)

[![Project Status: Active - The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)
[![codecov](https://codecov.io/gh/GuangchuangYu/DOSE/branch/master/graph/badge.svg)](https://codecov.io/gh/GuangchuangYu/DOSE/)
[![Last-changedate](https://img.shields.io/badge/last%20change-`r gsub('-', '--', Sys.Date())`-green.svg)](https://github.com/GuangchuangYu/DOSE/commits/master)
[![commit](http://www.bioconductor.org/shields/commits/bioc/DOSE.svg)](https://www.bioconductor.org/packages/devel/bioc/html/DOSE.html#svn_source)
[![GitHub forks](https://img.shields.io/github/forks/GuangchuangYu/DOSE.svg)](https://github.com/GuangchuangYu/DOSE/network) 
[![GitHub stars](https://img.shields.io/github/stars/GuangchuangYu/DOSE.svg)](https://github.com/GuangchuangYu/DOSE/stargazers) 


[![releaseVersion](https://img.shields.io/badge/release%20version-`r check_bioc("DOSE")$latest_version`-green.svg?style=flat)](https://bioconductor.org/packages/DOSE)
[![develVersion](https://img.shields.io/badge/devel%20version-`r get_devel_version("GuangchuangYu/DOSE")`-green.svg?style=flat)](https://github.com/GuangchuangYu/DOSE)
[![Bioc](http://www.bioconductor.org/shields/years-in-bioc/DOSE.svg)](https://www.bioconductor.org/packages/devel/bioc/html/DOSE.html#since)
[![post](http://www.bioconductor.org/shields/posts/DOSE.svg)](https://support.bioconductor.org/t/DOSE/)
[![download](http://www.bioconductor.org/shields/downloads/DOSE.svg)](https://bioconductor.org/packages/stats/bioc/DOSE/)



This package implements five methods proposed by *Resnik*, *Schlicker*, *Jiang*, *Lin* and *Wang* respectively for measuring semantic similarities among DO terms and gene products. Enrichment analyses including hypergeometric model and gene set enrichment analysis are also implemented for discovering disease associations of high-throughput biological data. 


[![Twitter](https://img.shields.io/twitter/url/https/github.com/GuangchuangYu/DOSE.svg?style=social)](https://twitter.com/intent/tweet?hashtags=DOSE&url=http://bioinformatics.oxfordjournals.org/content/31/4/608)

------------------------------------------------------------------------

Please cite the following article when using `DOSE`: 

__*G Yu*__, LG Wang, GR Yan, QY He. DOSE: an R/Bioconductor package for Disease Ontology Semantic and Enrichment analysis. __*Bioinformatics*__ 2015, 31(4):608-609. 

[![doi](https://img.shields.io/badge/doi-10.1093/bioinformatics/btu684-green.svg?style=flat)](http://dx.doi.org/10.1093/bioinformatics/btu684)
[![citation](https://img.shields.io/badge/cited%20by-`r sum(get_article_cite_history('DO5oG40AAAAJ', 'Ug5p-4gJ2f0C')$cites)`-green.svg?style=flat)](https://scholar.google.com.hk/scholar?oi=bibs&hl=en&cites=16627502277303919270)
[![Altmetric](https://img.shields.io/badge/Altmetric-`r x <- readLines("https://www.altmetric.com/details/2788597"); gsub("^.*score=(\\d+)\\D+.*$", '\\1', x[grep("style=donut&score=", x)])`-green.svg?style=flat)](https://www.altmetric.com/details/2788597)


------------------------------------------------------------------------

For details, please visit our project website, <https://guangchuangyu.github.io/DOSE>.

+ [Documentation](https://guangchuangyu.github.io/DOSE/documentation/)
+ [Featured Articles](https://guangchuangyu.github.io/DOSE/featuredArticles/)
+ [Feedback](https://guangchuangyu.github.io/DOSE/#feedback)

### Citation 

[![citation](https://img.shields.io/badge/cited%20by-`r sum(get_article_cite_history('DO5oG40AAAAJ', 'Ug5p-4gJ2f0C')$cites)`-green.svg?style=flat)](https://scholar.google.com.hk/scholar?oi=bibs&hl=en&cites=16627502277303919270)


```{r echo=F, comment=NA}
d <- get_article_cite_history('DO5oG40AAAAJ', 'Ug5p-4gJ2f0C')
txtplot(d$year, d$cites)
```

### Download stats

[![download](http://www.bioconductor.org/shields/downloads/DOSE.svg)](https://bioconductor.org/packages/stats/bioc/DOSE/)
[![total](https://img.shields.io/badge/downloads-`r sum(bioc_stats("DOSE")$Nb_of_downloads)`/total-blue.svg?style=flat)](https://bioconductor.org/packages/stats/bioc/DOSE/)
[![month](https://img.shields.io/badge/downloads-`r x <- bioc_stats("DOSE"); x$Nb_of_downloads[nrow(x)-1]`/month-blue.svg?style=flat)](https://bioconductor.org/packages/stats/bioc/DOSE/)

```{r echo=F, comment=NA}
x <- bioc_stats("DOSE")
x <- x[-nrow(x), ] # last month data is not complete
txtplot(Date2decimal(x$end), x$Nb_of_downloads, width=100, height=25)
```



