---
output:
  md_document:
    variant: gfm
html_preview: false
---


```{r echo=FALSE, results="hide", message=FALSE}
library("badger")
library("ypages")
```


# DOSE: Disease Ontology Semantic and Enrichment analysis

`r badge_bioc_release("DOSE", "green")`
`r badge_devel("guangchuangyu/DOSE", "green")`
[![Bioc](http://www.bioconductor.org/shields/years-in-bioc/DOSE.svg)](https://www.bioconductor.org/packages/devel/bioc/html/DOSE.html#since)
[![codecov](https://codecov.io/gh/GuangchuangYu/DOSE/branch/master/graph/badge.svg)](https://codecov.io/gh/GuangchuangYu/DOSE/)



[![Project Status: Active - The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)
[![platform](http://www.bioconductor.org/shields/availability/devel/DOSE.svg)](https://www.bioconductor.org/packages/devel/bioc/html/DOSE.html#archives)
[![Build Status](http://www.bioconductor.org/shields/build/devel/bioc/DOSE.svg)](https://bioconductor.org/checkResults/devel/bioc-LATEST/DOSE/)
[![Linux/Mac Travis Build Status](https://img.shields.io/travis/GuangchuangYu/DOSE/master.svg?label=Mac%20OSX%20%26%20Linux)](https://travis-ci.org/GuangchuangYu/DOSE)
[![AppVeyor Build Status](https://img.shields.io/appveyor/ci/Guangchuangyu/DOSE/master.svg?label=Windows)](https://ci.appveyor.com/project/GuangchuangYu/DOSE)



```{r comment="", echo=FALSE, results='asis'}
cat(packageDescription('DOSE')$Description)
```



For details, please visit <https://yulab-smu.top/biomedical-knowledge-mining-book/>.


## :writing_hand: Authors

Guangchuang YU  <https://yulab-smu.top>

School of Basic Medical Sciences, Southern Medical University


[![Twitter](https://img.shields.io/twitter/url/http/shields.io.svg?style=social&logo=twitter)](https://twitter.com/intent/tweet?hashtags=clusterProfiler&url=http://online.liebertpub.com/doi/abs/10.1089/omi.2011.0118&screen_name=guangchuangyu)
[![saythanks](https://img.shields.io/badge/say-thanks-ff69b4.svg)](https://saythanks.io/to/GuangchuangYu)
`r badger::badge_custom('follow me on', 'WeChat', 'green', 'https://guangchuangyu.github.io/blog_images/biobabble.jpg')`


------------------------------------------------------------------------

Please cite the following article when using `DOSE`:

__*G Yu*__, LG Wang, GR Yan, QY He. DOSE: an R/Bioconductor package for Disease Ontology Semantic and Enrichment analysis. __*Bioinformatics*__ 2015, 31(4):608-609.

`r badge_doi("10.1093/bioinformatics/btu684", "green")`
`r badge_altmetric("2788597", "green")`

<!--

------------------------------------------------------------------------





### Download stats

r badge_download_bioc("DOSE")`
r badge_bioc_download("DOSE", "total", "blue")`
r badge_bioc_download("DOSE", "month", "blue")`

-->


