---
output:
  md_document:
    variant: markdown_github
html_preview: false
---


```{r echo=FALSE, results="hide", message=FALSE}
library("txtplot")
library("ypages")
```


# DOSE: Disease Ontology Semantic and Enrichment analysis

`r badge_release("DOSE", "green")`
`r badge_devel("DOSE", "green")`
[![Bioc](http://www.bioconductor.org/shields/years-in-bioc/DOSE.svg)](https://www.bioconductor.org/packages/devel/bioc/html/DOSE.html#since)
`r badge_download("DOSE", "total", "blue")`
`r badge_download("DOSE", "month", "blue")`


[![Project Status: Active - The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)
[![codecov](https://codecov.io/gh/GuangchuangYu/DOSE/branch/master/graph/badge.svg)](https://codecov.io/gh/GuangchuangYu/DOSE/)
[![Last-changedate](https://img.shields.io/badge/last%20change-`r gsub('-', '--', Sys.Date())`-green.svg)](https://github.com/GuangchuangYu/DOSE/commits/master)
[![commit](http://www.bioconductor.org/shields/commits/bioc/DOSE.svg)](https://www.bioconductor.org/packages/devel/bioc/html/DOSE.html#svn_source)
[![GitHub forks](https://img.shields.io/github/forks/GuangchuangYu/DOSE.svg)](https://github.com/GuangchuangYu/DOSE/network)
[![GitHub stars](https://img.shields.io/github/stars/GuangchuangYu/DOSE.svg)](https://github.com/GuangchuangYu/DOSE/stargazers)


[![platform](http://www.bioconductor.org/shields/availability/devel/DOSE.svg)](https://www.bioconductor.org/packages/devel/bioc/html/DOSE.html#archives)
[![Build Status](http://www.bioconductor.org/shields/build/devel/bioc/DOSE.svg)](https://bioconductor.org/checkResults/devel/bioc-LATEST/DOSE/)
[![Linux/Mac Travis Build Status](https://img.shields.io/travis/GuangchuangYu/DOSE/master.svg?label=Mac%20OSX%20%26%20Linux)](https://travis-ci.org/GuangchuangYu/DOSE)
[![AppVeyor Build Status](https://img.shields.io/appveyor/ci/Guangchuangyu/DOSE/master.svg?label=Windows)](https://ci.appveyor.com/project/GuangchuangYu/DOSE)
[![install with bioconda](https://img.shields.io/badge/install%20with-bioconda-green.svg?style=flat)](http://bioconda.github.io/recipes/bioconductor-dose/README.html)



This package implements five methods proposed by *Resnik*, *Schlicker*, *Jiang*, *Lin* and *Wang* respectively for measuring semantic similarities among DO terms and gene products. Enrichment analyses including hypergeometric model and gene set enrichment analysis are also implemented for discovering disease associations of high-throughput biological data.


[![Twitter](https://img.shields.io/twitter/url/https/github.com/GuangchuangYu/DOSE.svg?style=social)](https://twitter.com/intent/tweet?hashtags=DOSE&url=http://bioinformatics.oxfordjournals.org/content/31/4/608)

------------------------------------------------------------------------

Please cite the following article when using `DOSE`:

__*G Yu*__, LG Wang, GR Yan, QY He. DOSE: an R/Bioconductor package for Disease Ontology Semantic and Enrichment analysis. __*Bioinformatics*__ 2015, 31(4):608-609.

`r badge_doi("10.1093/bioinformatics/btu684", "green")`
`r badge_citation("Ug5p-4gJ2f0C", "16627502277303919270", "green")`
`r badge_altmetric("2788597", "green")`


------------------------------------------------------------------------

For details, please visit our project website, <https://guangchuangyu.github.io/DOSE>.

+ [Documentation](https://guangchuangyu.github.io/DOSE/documentation/)
+ [Featured Articles](https://guangchuangyu.github.io/DOSE/featuredArticles/)
+ [Feedback](https://guangchuangyu.github.io/DOSE/#feedback)

### Citation

`r badge_citation("Ug5p-4gJ2f0C", "16627502277303919270", "green")`


```{r echo=F, comment=NA}
d <- get_article_cite_history('DO5oG40AAAAJ', 'Ug5p-4gJ2f0C')
txtplot(d$year, d$cites)
```

### Download stats

[![download](http://www.bioconductor.org/shields/downloads/DOSE.svg)](https://bioconductor.org/packages/stats/bioc/DOSE/)
`r badge_download("DOSE", "total", "blue")`
`r badge_download("DOSE", "month", "blue")`

```{r echo=F, comment=NA}
x <- dlstats::bioc_stats("DOSE")
x <- x[-nrow(x), ] # last month data is not complete
txtplot(ggtree::Date2decimal(x$end), x$Nb_of_downloads, width=100, height=25)
```



