---
title: "DOSE: Disease Ontology Semantic and Enrichment analysis"
author: "Guangchuang Yu\\

        School of Basic Medical Sciences, Southern Medical University"
date: "`r Sys.Date()`"
bibliography: DOSE.bib
biblio-style: apalike
output:
  prettydoc::html_pretty:
    toc: true
    theme: cayman
    highlight: github
  pdf_document:
    toc: true
vignette: >
  %\VignetteIndexEntry{DOSE}
  %\VignettePackage{DOSE}
  % \VignetteEngine{knitr::rmarkdown}
  % \usepackage[utf8]{inputenc}
  %\VignetteEncoding{UTF-8}
---

```{r style, echo=FALSE, results="asis", message=FALSE}
knitr::opts_chunk$set(tidy = FALSE,
                      warning = FALSE,
                      message = FALSE)

CRANpkg <- function (pkg) {
    cran <- "https://CRAN.R-project.org/package"
    fmt <- "[%s](%s=%s)"
    sprintf(fmt, pkg, cran, pkg)
}

Biocpkg <- function (pkg) {
    sprintf("[%s](http://bioconductor.org/packages/%s)", pkg, pkg)
}

```


```{r echo=FALSE, results='hide', message=FALSE}
library(DOSE)
```

Disease Ontology (DO)[@schriml_disease_2011] aims to provide an open source ontology for the integration of biomedical data that is associated with human disease.
We developed `r Biocpkg("DOSE")`[@yu_dose_2015] package to promote the investigation of diseases. `r Biocpkg("DOSE")` provides five methods including _Resnik_, _Lin_, _Jiang_, _Rel_ and _Wang_ for measuring semantic similarities among DO terms and gene products; Hypergeometric model and Gene Set Enrichment Analysis (GSEA) were also implemented for associating disease with gene list and extracting disease association insight from genome wide expression profiles.

The enrichment analyses support Disease Ontology (DO)[@schriml_disease_2011], Network of Cancer Gene (NCG)[@omer_ncg] and DisGeNET[@janet_disgenet]. In addition, several visualization methods were provided by `r Biocpkg("enrichplot")` to help interpreting semantic and enrichment results.


# Citation

If you use `r Biocpkg("DOSE")` in published research, please cite G. Yu (2015).


__*G Yu*__, LG Wang, GR Yan, QY He. DOSE: an R/Bioconductor package for Disease Ontology Semantic and Enrichment analysis. __*Bioinformatics*__ 2015, 31(4):608-609. <http://dx.doi.org/10.1093/bioinformatics/btu684>.


# Vignette 


Please go to <https://yulab-smu.github.io/clusterProfiler-book> for the full vignette of enrichment analysis.

For semantic analysis, please go to <https://bioconductor.org/packages/devel/bioc/vignettes/DOSE/inst/doc/semanticAnalysis.html>.



# Need helps?


If you have questions/issues, please visit
[DOSE homepage](https://guangchuangyu.github.io/software/DOSE/) first.
Your problems are mostly documented. If you think you found a bug, please follow
[the guide](https://guangchuangyu.github.io/2016/07/how-to-bug-author/) and
provide a reproducible example to be posted
on
[github issue tracker](https://github.com/GuangchuangYu/DOSE/issues).
For questions, please post
to [Bioconductor support site](https://support.bioconductor.org/) and tag your
post with *DOSE*.



# References

